---
title: "Sustainability of Brazilian forest concessions"
output: 
  bookdown::word_document2:
  fig_caption: yes
  keep_md: true
bibliography: myBiblio.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, results='hide', 
                      root.dir = rprojroot::find_rstudio_root_file())
# load libraries
packages_needed <- c("knitr", "data.table", "ggplot2", "parallel", "sp", "truncnorm", "tmap")
packages_to_install <- packages_needed[!( packages_needed %in% rownames(installed.packages()))]
if (length(packages_to_install) > 0)
  install.packages(packages_to_install)
lapply(packages_needed, require, character.only = TRUE)

## source all functions
file.sources = list.files(path = "R/", pattern="*.R", full.names = TRUE)
sapply(file.sources, source, .GlobalEnv)

if (!dir.exists("cache"))
  dir.create("cache")
```

# Introduction

Legal logging in the tropics is typically governed by simple rules that set a minimum cutting diameter and a minimum cutting cycle of 25-35 years [@Putz2008]. Timber harvests in diverse forests are selective because few species are marketable and few trees meet the size and bole quality requirements. In the Amazon, selective logging regulations typically set a rotation cycle of 20 to 35 years with a logging intensity varying from 15 to 30 m$^3$ of harvested timber per ha. Such rules, promulgated by governments since the middle of last century, are generally based on an assumed post-logging rate of commercial timber volume increments of about 1 m$^3$.ha$^{-1}$.year$^{-1}$ (0.86 m$^3$.ha$^{-1}$.year$^{-1}$ in Brazilian Amazon). These rules are set to accommodate processing technologies and market demands, rather than the biology and conservation of the harvested species [@Sist2008]. Unfortunately for the goal of sustainability, research reports over the past decades demonstrated that volume recovery rates fall short of this expectation by 50% [reviewed by @Putz2012]. A recent simulation of post-logging timber volume recovery rates in the Amazon Basin showed that even with cutting cycles of 65 years and logging intensities of only 20 m$^3$.ha$^{-1}$, logged forests recover only 70% of their pre-logging timber stocks [@Piponiot2019a]. Other researchers showed that current harvest regimes can only be sustained over multiple cycles if high-value slow-growing hardwoods are replaced by fast-growing species with low density wood of little market value [@Alder2000;@Keller2004;@Phillips2004;@Gardingen2006;@Sist2007;@Schulze2008]. Here we use a timber recovery model [@Piponiot2019a] to estimate the timber volumes that could be produced by all the logging concessions in the Brazilian Amazon with different cutting cycle lengths, logging intensities, and lengths of the list of commercial species. 

Despite low rates of post-harvest timber stock recovery, selective logging is still widespread and economically important land use in the tropics in general and Amazon Basin in particular. In that region, forest degradation due to illegal logging is widespread [@Brancalion2018;@Finer2014;@Potapov2017] and, in the Brazilian Amazon, affects now bigger areas than deforestation [@Matriciardi2020]. This makes difficult to assess the region's current potential for timber production. What is clear, in contrast, is that without control of illegal logging and improved practices where logging is legal, timber yields from logged forests will decline dramatically [@Putz2012;@Piponiot2019a], decreasing the likelihood of their meeting the demand for timber.

In 2006, the Brazilian Forest Service (SFB) was created to establish a very ambitious system of long-term logging concessions [@Brazil2006]. The goals are to provide a legal framework for sustainable timber production in Amazonian forests while reducing illegal logging. Forest concessions in the Brazilian Amazon currently cover only 1.6 million ha [@SFB2019], but the maximum potential area is estimated up to 60 million ha [@Bomfim2016]. The current timber production rate from established forest concessions is 221,000 m$^3$ per year, which is only 2% of the timber extracted from the region [@SFB2019]. Given that these concessions are to be managed with a 50 cm minimum cutting diameter (with the exception of *Swientenia macrophylla*: 60 cm) and a 25-35 year cutting cycle coupled with rising demand for wood products, the assessment of the expected timber production from these forests over the long term is warranted. 

In this paper we assess the timber yield sustainability of different logging intensities and harvest cycle durations from forest concessions in the Brazilian Amazon. We assess whether the potential annual timber production is adequate to meet the estimated present timber demand of 11 Mm$^3$.yr$^{-1}$ [@Vidal2020;@SFB2019].

# Methods

## Study areas - Brazilian concessions

```{r get-prop-concess}
if (!file.exists("data/pconcessions.Rdata")) {
  ## current concessions ####
  concessions <- rgdal::readOGR(dsn = "data_sfb/final", layer = "BRconc")
  # define new raster with resolution = 1 degree
  ext2 <- floor(raster::extent(concessions)) + c(0, 1, 0, 1)
  rst <- raster::raster(resolution = 0.01, ext = ext2)
  ## transform shapefile into raster with 1 degree resolution
  raster_concessions <- raster::rasterize(concessions, rst)
  raster_concessions2 <- !is.na(raster_concessions)
  pconcessions <- raster::aggregate(raster_concessions2, fact = 100)
  pconcessions <- raster::as.data.frame(pconcessions, xy = TRUE)
  colnames(pconcessions) <- c("long", "lat", "pConcess")
  
  ## all potential concessions ####
  flonas <- rgdal::readOGR(dsn = "data_sfb/final", layer = "BRflopub")
  # define new raster with resolution = 1 degree
  ext2 <- floor(raster::extent(flonas)) + c(0, 1, 0, 1)
  rst <- raster::raster(resolution = 0.01, ext = ext2)
  raster_flonas <- raster::rasterize(flonas, rst)
  raster_flonas2 <- !is.na(raster_flonas)
  pflonas <- raster::aggregate(raster_flonas2, fact = 100)
  pflonas <- raster::as.data.frame(pflonas, xy = TRUE)
  colnames(pflonas) <- c("long", "lat", "pFlona")
  
  pconcessions <- merge(pconcessions, pflonas, by = c("long", "lat"), all = TRUE)
  pconcessions[is.na(pconcessions$pConcess), "pConcess"] <- 0
  pconcessions <- subset(pconcessions, pConcess > 0 | pFlona > 0)
  save(pconcessions, file = "data/pconcessions.Rdata")
} else load("data/pconcessions.Rdata")
```

Our study focuses on forest concessions in the Brazilian Amazon (Fig. \@ref(fig:map-brazil-concess)). These concessions are located in public forests and currently cover 1.6 Mha, of which 1.05 Mha are managed by the SFB, and 0.6 Mha are managed by state-level agencies [@SFB2019]. We defined the area of all potential concessions as the area of all public forests that (i) are in the Brazilian Amazon biome, (ii) are designated for sustainable use, and (iii) are not community forests - although community forest management is legal and currently cover around 260.000 ha [@Miranda2020], indigenous territories or military areas (as defined in @SFB2019, p. 112; Fig. \@ref(fig:map-brazil-concess)]. Based on this definition, the potential concession area in the Brazilian Amazon cover an estimated of 35 Mha. This area is lower than the 60 Mha estimated by @Bomfim2016; xxx explanation. 

```{r map-brazil-concess, fig.height = 5, fig.width = 8, fig.cap = "Forest concessions in the Brazilian Amazon. Current federal concessions are in red; potential concessions (public forests designated for sustainable use) are in blue [retrieved from Brazilian Forest Service and IDEFLOR websites [@SFB2020;@SFB2019a;@IDEFLOR2020]]."}
concessions <- rgdal::readOGR(dsn = "data_sfb/final", layer = "BRconc")
concessions$ppf <- "current"
flopub <- rgdal::readOGR(dsn = "data_sfb/final", layer = "BRflopub")
flopub <- rbind(flopub, concessions)
map.samerica <- rgdal::readOGR("D:/Mes Donnees/maps/borders/South_America", "South_America")
map.samerica <- raster::crop(map.samerica, raster::extent(-85,-33,-56,13))
nb_region = sf::st_as_sfc(sf::st_bbox(c(xmin = -74, xmax = -44, ymin = -11, ymax = 5),
                    crs = sf::st_crs(map.samerica)))
conc_map = tm_shape(map.samerica, bbox = nb_region) +  ## maybe just subset world map
  tm_polygons(alpha = 0) +
  tm_shape(flopub) +
  tm_fill(col = "ppf", 
          title = "Concession area", 
          palette = c("firebrick3", "dodgerblue2")) + 
  tm_compass(type = "8star", position = c("right", "bottom")) +
  tm_scale_bar(breaks = c(0, 100, 200), text.size = 1)
sam_map = tm_shape(map.samerica) + tm_polygons() + 
  tm_shape(nb_region) + tm_borders(lwd = 3) 

conc_map
print(sam_map, vp = grid::viewport(0.1, 0.7, width = 0.2, height = 0.4))
``` 


```{r get-area-pixels}
raster_conc <- expand.grid(long = seq(min(pconcessions$long), max(pconcessions$long)), 
                           lat = seq(min(pconcessions$lat), max(pconcessions$lat)))
coordinates(raster_conc) <- ~long + lat
proj4string(raster_conc) <- CRS("+proj=longlat +datum=WGS84")
gridded(raster_conc) <- TRUE
areatot <- raster::area(raster::raster(raster_conc)) ## pixel area, in km2
raster_conc$area <- raster::extract(areatot, raster_conc)*100 ## km -> ha

pconcessions <- merge(pconcessions, 
                      as.data.frame(raster_conc), 
                      by = c("long", "lat"))

## all pixels with flonas (potential concessions): will be used in simulations
locations <- subset(pconcessions, pFlona > 0 | pConcess > 0)
locations$site <- 1:nrow(locations)
## multiply by factor pi = 8.2/(8.2+5.9)
## TODO add this to iterations as pi = rbeta(100, 8.2, 5.9)
pi = 8.2/(8.2+5.9)
locations$pConcess <- locations$pConcess*pi
locations$pFlona <- locations$pFlona*pi
```


## The VDDE model 

In this study we used the volume dynamics with differential equations model (VDDE) [@Piponiot2018]. The VDDE model calculates the volume of all live trees $\geq$ 50 cm diameter at breast height (DBH), the standard minimum cutting size in the Brazilian Amazon. The portion of this volume composed of commercial species is referred to as the commercial volume.

The input variables set for each simulation are logging intensity, logging cycle length, and the initial proportion of the volume that is commercial ($\omega_0$). Low values of $\omega_0$ represent highly selective logging where only the most valuable species are harvested whereas high values mean that most species are logged. Other parameters of the VDDE model are spatially defined at a 1$^\circ$ resolution. The model was calibrated for the Amazon Basin with a Bayesian framework with data from 3500 ha of forest plots, among which 845 ha are from 15 sites monitored for as long as 30 years after selective logging [@Sist2015;@Piponiot2019a]. Predictions of commercial volume recovery rates thereby vary with location.

Each logging cycle includes the harvest itself as a function of logging intensity and forest characteristics and the post-logging volume recovery phase, which varies with logging cycle length and forest characteristics. Logging lowers both the total volume and the proportion of commercial volume, but both then increase during the recovery phase, although the proportion of commercial volume takes longer to recover because it relies solely on the recruitment of trees < 50 cm DBH [@Piponiot2019a]. These two steps are sequentially repeated to simulate 1000 years of logging.

The results are then multiplied by the area of current or potential concessions in each 1$^\circ$ pixel, and by a factor $\pi \sim \mathcal{Beta}(8.2, 5.9)$ between 0 and 1, with a mean value of 58%. This factor $\pi$, which was calibrated with data from logging concessions in French Guiana, reflects the ratio between logged areas and the initially allocated areas, mostly because of slope restrictions and riparian reserves, but also heavy forest degradation because of illegal logging and other disturbances [@Verissimo2006;@Piponiot2019a]. 

Uncertainties are propagated throughout the model by drawing all parameter values from their calibrated distribution [from @Piponiot2019a], and simulating logging cycles with these parameter values. This process is repeated 100 times and summary statistics (medians and 95% credibility intervals) are calculated at each time step.

## Testing scenarios

We tested 27 different scenarios by using combinations of the following inputs: (i) initial proportion of commercial volume: 20% (highly selective), 50% (intermediate) or 90% (non-selective); (ii) logging intensity: 10 m3.ha-1 (low), 20 m$^3$.ha$^{-1}$ (intermediate) or 30 m$^3$.ha$^{-1}$ (high); (iii) cutting cycle length: 20 years (short), 35 years (intermediate) or 60 years (long). These scenarios are then applied to the current area of concessions and to the area of all potential concessions (see “Study areas”).

For each scenario we simulated 1000 years of logging cycles, and determined the duration of maintained timber production, i.e. the time before timber stocks become insufficient to maintain a constant timber production as illustrated in Fig. \@ref(fig:illust-sust-log). This maintained production is different from sustained timber production which theoretically shows a constant timber yield and stock over time (Fig. \@ref(fig:illust-sust-log)).

```{r illust-sust-log, fig.height = 4, fig.width = 8, fig.cap = "Illustration of the duration of maintained and sustained timber production. The x-axis represents years after the first selective harvest, and the y-axis represents commercial volumes as simulated by the model with a logging intensity of 10 m$^3$.ha$^{-1}$ and a logging cycle of 60 years. At each harvest, commercial volumes decrease (blue segments). If logging cycles are not long enough to allow recovery, the commercial volume decreases until it is not sufficient to maintain a constant production (10, 20 or 30 m$^3$.ha$^{-1}$, red segments). The time taken to reach this limit is the duration of the maintained production. In the sustained timber production, both timber yield and stocks remain constant."}
logCyc = 60
logInt = 10

dfsimu = simulateLogging(
  timeLength = 1000,
  logIntensity = logInt,
  logCycle = logCyc,
  omega0 = 0.5,
  uncertainties = FALSE,
  longitude = -60,
  latitude = 0
)
dfvol = dcast(subset(dfsimu, variable!="vextReal"), ncycle + trec ~ variable)
dfvol$vcom = dfvol$volume*dfvol$omega
dfcycle = dcast(dfvol, ncycle ~ trec, value.var = "vcom")
colnames(dfcycle) <- c("ncycle", "pl", "pr")
dfcycle[, pl := c(pl[-1], NA)]
dfvext = subset(dfsimu, variable=="vextReal")
tsust = max(dfvext[value==logInt, "ncycle"])*logCyc

## 'sustainable logging'
slope = (dfvol[ncycle==2&trec==60]$vcom-dfvol[ncycle==2&trec==1]$vcom)/59
interc = dfvol[ncycle==2&trec==1]$vcom - 61*slope
lsust = (dfvol[ncycle==1&trec==60]$vcom-interc)/slope - 61

dfsust = data.frame(x = lsust*c(0, rep(1:5, each = 2)) + 61 + rep(1:5,each = 2)[1:11], 
                    y = rep(c(dfvol[ncycle==2&trec==1]$vcom, 
                              dfvol[ncycle==1&trec==60]$vcom), 8)[1:11])
  
ggplot() +
  geom_line(data = dfsust, aes(x=x, y=y)) +
  geom_vline(xintercept = tsust, lty = 2) + 
  geom_line(data = dfvol, aes(x = trec+logCyc*(ncycle-1), y = vcom)) + 
  geom_segment(data = dfcycle, 
               aes(x = logCyc*ncycle, xend = logCyc*ncycle+1, 
                   yend = pl, y = pr), col = 4, lwd = 1) + 
  geom_segment(data = subset(dfcycle, ncycle >= max(dfvext[value==logInt, "ncycle"])), 
               aes(x = logCyc*ncycle, xend = logCyc*ncycle+1, 
                   yend = pl, y = pr), col = 2, lwd = 1) + 
  annotate(geom = "segment", x = 0, xend = tsust, y = 57, yend = 57,
           arrow = arrow(length = unit(0.2, "cm"), ends = "both")) +
  annotate(geom = "text", x = c(500, 820, 820, 350), y = c(25, 10, 45, 59), 
           col = c(4,2, 1, 1),
           label = c("Maintained\nproduction", "Timber depletion", 
                     "Sustained timber\nproduction",
                     "Duration of maintained production (tsust)")) +
  labs(x = "Time after first logging event (yr)", y = "Commercial volume") + 
  theme_classic() + 
  scale_x_continuous(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0)) +
  expand_limits(y = c(0,60))
``` 

```{r define-simulations}
timeLength <- 200
omega0 <- c(0.2, 0.5, 0.9)
logIntensity <- (1:3)*10
logCycle <- c(20, 35, 60)
simulations <- expand.grid(omega0 = omega0, 
                           logIntensity = logIntensity, 
                           logCycle = logCycle)
simulations$simid <- 1:nrow(simulations)
```

```{r run-simulations}
if (!dir.exists("cache/simulations_concessionsBR/"))
  dir.create("cache/simulations_concessionsBR/")

all_i = 1:nrow(simulations)
del_i = as.numeric(gsub("recov|\\.Rdata", "", list.files("cache/simulations_concessionsBR/")))
missing_simus <- all_i[!all_i%in%del_i]

if (length(missing_simus) > 0) {
  source("R/simulateLogging.R")
  
  # Calculate the number of cores
  no_cores <- detectCores() - 2
  # Initiate cluster
  cl <- makeCluster(no_cores)
  clusterEvalQ(cl, library(data.table))
  clusterExport(cl, varlist = c("simulateLogging", "simulations", "timeLength", "locations", "rtruncnorm"))
  
  recovCycles <- parLapply(cl, missing_simus, function(j) {
    recov <- simulateLogging(
      timeLength = 1000,
      logIntensity = simulations$logIntensity[j],
      logCycle = simulations$logCycle[j],
      omega0 = simulations$omega0[j],
      longitude = locations$long,
      latitude = locations$lat,
      uncertainties = TRUE, keepAll = FALSE
    )
    save(recov, file = paste0("cache/simulations_concessionsBR/recov", j, ".Rdata"))
    return(recov)
  })
  
  stopCluster(cl)
}
```

```{r prepare-results}
outputs <- lapply(list.files("cache/simulations_concessionsBR/", full.name = TRUE), function(fnm){
  load(fnm)
  get("recov")
})
names(outputs) <- gsub("recov|\\.Rdata", "", list.files("cache/simulations_concessionsBR/"))

results <- rbindlist(outputs, idcol = "simid")
results$simid <- as.numeric(results$simid)

## multiply by total area ####
results <- merge(results, locations, by = "site")
results <- melt(results, measure.vars = c("pConcess", "pFlona"), 
                variable.name = "ppf", value.name = "pArea", value.factor = TRUE)
levels(results$ppf) <- c("current", "potential")
results <- subset(results, pArea > 0)
```

```{r get-timber-production}
dfvext <- subset(results, variable == "vextReal")
## multiply by area and sum (or mean for omega0)
dfvext <- dfvext[, .(lwr = sum(lwr*pArea*area),
                     med = sum(med*pArea*area), 
                     upr = sum(upr*pArea*area)), 
                 .(simid, ncycle, trec, ppf)]
# add simulation info
dfvext <- merge(dfvext, simulations, by = "simid")
```

```{r get-timber-stocks}
dfvolume <- subset(results, variable != "vextReal")
dfvolume <-
  dcast(dfvolume, value.var = c("lwr", "med", "upr"),
        ncycle + trec + simid + ppf + area + pArea + site ~ variable
  )
dfvolume$lwr <- dfvolume$lwr_omega*dfvolume$lwr_volume
dfvolume$med <- dfvolume$med_omega*dfvolume$med_volume
dfvolume$upr <- dfvolume$upr_omega*dfvolume$upr_volume
dfvolume <- dfvolume[, .(lwr = sum(lwr*pArea*area),
                         med = sum(med*pArea*area), 
                         upr = sum(upr*pArea*area)), 
                     .(simid, ncycle, trec, ppf)]
dfvolume <- dfvolume[, c("ncycle", "trec", "simid", "ppf", "lwr", "med", "upr")]
```

```{r add-volume0}
## add volume0
## initial maturity and volume
load("data/paramStan.Rdata")
dfPred <- spatialPars(locations[, c("long", "lat")], uncert = TRUE)
dfPred$dti <- sample(pars_dti, nrow(dfPred), replace = TRUE)

dfPred[, matInit := stem_mort ^ (-lambda_ti) * (1 - dti)]
dfPred[, vol0 := volume(matInit, aG_FORMIND, aM, bP, bM, theta)]

dfvol0 <- dfPred[, .(variable = "volume", 
                     lwr = quantile(vol0, 0.025), 
                     med = quantile(vol0, 0.5), 
                     upr = quantile(vol0, 0.975)), 
                 .(long, lat)]
dfvol0 <- merge(dfvol0, locations, by = c("long", "lat"))
dfvol0 <- melt(dfvol0, measure.vars = c("pConcess", "pFlona"), 
               variable.name = "ppf", value.name = "pArea", value.factor = TRUE)
levels(dfvol0$ppf) <- c("current", "potential")
dfvol0 <- subset(dfvol0, pArea > 0)
dfvol0 <- dfvol0[, .(lwr = sum(lwr*pArea*area), 
                     med = sum(med*pArea*area),
                     upr = sum(upr*pArea*area),
                     trec = 0, ncycle = 1), .(ppf)]

dfres <- expand.grid(ppf = c("current", "potential"), simid = simulations$simid)
dfres <- merge(dfres, simulations[, c("simid", "omega0")])
dfres <- merge(dfvol0, dfres, by = "ppf")
dfres[, `:=`(lwr = omega0*lwr, med = omega0*med, upr = omega0*upr)]
dfvolume <- rbind(dfvolume, dfres[, colnames(dfvolume), with = FALSE])
```

```{r annualize-volume}
dfvolume <- subset(dfvolume, trec != 1)
dfvolume <- merge(dfvolume, simulations, by = "simid")
dfvolume[, t := (ncycle-1)*logCycle + trec]
```



# Results

```{r calc-tsust}
dftsust <- lapply(outputs, function(df) { 
  df <- subset(df, variable == "vextReal")
  ## add site information
  df <- merge(df, locations, by = "site")
  df <- melt(df, measure.vars = c("pConcess", "pFlona"), 
             variable.name = "ppf", value.name = "pArea", 
             value.factor = TRUE)
  levels(df$ppf) <- c("current", "potential")
  df <- subset(df, pArea > 0)
  df[, `:=`(med = med*area*pArea, 
            lwr = lwr*area*pArea, 
            upr = upr*area*pArea)]
  df <- df[, .(lwr = sum(lwr), med = sum(med), upr = sum(upr)),
           .(ppf, ncycle)]
  ini <- df[ncycle==1, -"ncycle"]
  ini$variable <- "prodi"
  res <- rbind(ini, df[, .(lwr = max(ncycle[lwr>=0.95*lwr[ncycle==1]]), 
                           med = max(ncycle[med>=0.95*med[ncycle==1]]), 
                           upr = max(ncycle[upr>=0.95*upr[ncycle==1]]), 
                           variable="tsust"), .(ppf)])
  return(res)
})
dftsust <- rbindlist(dftsust, idcol = "simid")
dftsust$simid <- as.numeric(dftsust$simid)
dftsust <- merge(dftsust, simulations, by = "simid")
dftsust[variable == "prodi", `:=`(lwr = lwr/logCycle, 
                                  med = med/logCycle, 
                                  upr = upr/logCycle)]
dftsust[variable == "tsust", `:=`(lwr = lwr*logCycle, 
                                  med = med*logCycle, 
                                  upr = upr*logCycle)]
dftsust[variable == "tsust" & med > 1000, med := 1000]
dftsust[variable == "tsust" & upr > 1000, upr := 1000]
dftsust = dcast(dftsust, omega0 + logIntensity + logCycle + ppf ~ variable, 
                value.var = c("lwr", "med", "upr"))
```

```{r tsust-vs-prodi, fig.width = 10, fig.height = 8, fig.cap = "Tradeoff between timber production and sustainability. The x-axis is the annual timber production under each scenario, and in all areas considered in the scenario (left panels: current concessions; right panels: potential concessions). The y-axis is the duration of constant production in each scenario, in years. The points are the median value over all simulations for each scenario; the vertical and horizontal error bars are the 95% credibility intervals. Colors represent logging rules (3 logging intensities x 3 logging cycle lengths) and the 3 values of initial proportion of commercial volume (omega_0) are represented by different panels, in increasing order from top to bottom. The target production of timber is 11 Mm$^3$.yr$^{-1}$, corresponding to the current timber production in Brazilian Amazonian forests. Only a few scenarios in the right panels (all potential concessions) are above this target, and all of them have a median duration of constant production lower than 200 years."}
## legend
palette <- function (name, indices = c(7,5,3)) {
  RColorBrewer::brewer.pal(9, name)[indices]
}
colours <- c(as.vector(sapply(c("Blues", "Purples","Reds"), palette)))
colour_palette <- c("LS"= colours[1], "LM"=colours[2], "LL"=colours[3],
                    "MS"=colours[4], "MM"=colours[5], "ML"=colours[6],
                    "HS"=colours[7], "HM"=colours[8], "HL"=colours[9])
df_rules <- data.table(logRules = c("LS","LM","LL","MS","MM","ML","HS","HM","HL"),
                       logIntensity = rep(10*1:3, each=3),
                       logCycle = rep(c(20,35,60), 3))
rules_legend <- ggplot(df_rules, aes(x=logIntensity, y=as.factor(logCycle),
                                     fill=logRules, label=logRules)) + 
  geom_raster() + scale_fill_manual(values = colour_palette) + 
  # geom_text(size=5) + 
  labs(x =expression("Logging intensity ("*m^3*ha^{-1}*")"), y = "Logging cycle (yrs)") + 
  theme(legend.position="none", aspect.ratio=1, 
        panel.background = element_blank(), 
        plot.margin = margin(2, 2, 2, 2, "cm"),
        text = element_text(size=10), 
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 15, r = 0, b = 0, l = 0))) 

## plot
dftsust <- merge(dftsust, df_rules, by = c("logIntensity", "logCycle"))
dftsust$ppf = factor(dftsust$ppf)
levels(dftsust$ppf) = c("Current concessions", "Potential concessions")
dftsust$omega0 = paste0("omega0 = ", dftsust$omega0*100, "%")
## objective production line only on 2nd facet
add_objProd <- data.frame(ppf="Potential concessions", pos = c(125, 700, 700),
                          omega0 = unique(dftsust$omega0),
                          obj = 11, text = "Target production")

graph <- ggplot(dftsust, aes(y=med_tsust, x = med_prodi/1e6)) +
  geom_vline(data = add_objProd, aes(xintercept = obj), lty=2) +
  geom_text(data = add_objProd, aes(x = obj+1, y = pos, label = text), angle = 90) +
  # ggrepel::geom_text_repel(aes(label = logRules), col = 1) +
  geom_pointrange(aes(ymin=lwr_tsust, ymax=upr_tsust, color = logRules)) +
  geom_errorbarh(height = 0, aes(xmin = lwr_prodi/1e6, xmax = upr_prodi/1e6, color = logRules)) +
  facet_grid(omega0 ~ ppf, scales = "free") +
  labs(x = bquote("Annual timber production (M"* m^3*.*yr^-1*")"), 
       y = "Duration of constant production (yrs)", 
       color = "Logging rules") + 
  scale_color_manual(values = colour_palette) +
  theme_bw() +
  theme(legend.position = "none") 

ggpubr::ggarrange(graph, rules_legend, widths = c(9,4))
```


```{r sust-production, results = "asis"}
dfsave = copy(dftsust)
dftsust[, lab_tsust := paste0(gsub("1000", "1000+", med_tsust), 
                              " yr (", lwr_tsust, "-", 
                              gsub("1000", "1000+", upr_tsust), ")")]
dftsust[, omega0 := gsub("omega0 = ", "", omega0)]
dftsust[, ppf := gsub(" concessions", "", ppf)]
dftsust[, logIntensity := paste(logIntensity, "m3/ha")]
dftsust[, logCycle := paste(logCycle, "yr")]
dftsust <- dcast(dftsust, omega0 + logIntensity + logCycle ~ ppf, value.var = "lab_tsust")

kable(dftsust[, c("omega0", "logIntensity", "logCycle", "Current", "Potential")],
      col.names = c("Comm. volume", "Log. intensity", "Log. cycle",
                    "Duration of constant production - Current concession area", "Duration of constant production - Potential concession area"),
      caption = "Sustainability of all 27 scenarios, characterized by the duration of constant timber production (yrs, last 2 columns). The first 3 columns correspond to the input variables: the proportion of commercial volume (%); logging intensity (m$^3$.ha${^-1}$); logging cycle length (yr). The 2 last columns are the duration of constant timber production in current concession areas and in all potential concessions, as the median value of all iterations, followed by the 95% credibility interval (between parentheses).")
```

```{r timber-stocks-br, fig.height=7, fig.width = 9, fig.cap="Commercial volume stocks in all potential concession areas for all sustainable scenarios."}
dfsave$simid <- as.numeric(as.factor(paste(dfsave$omega0, dfsave$logCycle, dfsave$logIntensity, dfsave$ppf)))
dfvolume$simid <- as.numeric(as.factor(paste(dfvolume$omega0, dfvolume$logCycle, dfvolume$logIntensity, dfvolume$ppf)))
sust_simid <- dfsave$simid[dfsave$med_tsust>=500]
dfvol_fig <- subset(dfvolume, simid %in% sust_simid)

dfvol_fig$logrules <- paste("Cycle:", dfvol_fig$logCycle, "yrs; Intensity:", 
                            dfvol_fig$logIntensity, "m3/ha; omega0:", 
                            as.numeric(dfvol_fig$omega0)*100,"%")

ggplot(subset(dfvol_fig, ppf=="potential"), aes(x=t, y = med/1e6)) +
  geom_line(aes(colour = logrules)) +
  geom_ribbon(alpha = 0.2, colour = NA,
              aes(fill = logrules,
                  ymin = lwr/1e6,
                  ymax = upr/1e6)) +
  # facet_wrap(~ ppf, scales = "free", nrow = 2)  +
  labs(x = "Time aftre the first logging event (yrs)", y = bquote("Commercial volume (M"*m^3*")"), 
       color = "Logging rules", fill = "Logging rules") +
  theme_classic() +
  theme(legend.position = "top") +
  scale_x_continuous(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0)) +
  guides(color = guide_legend(nrow=4))
ggsave("fig-sustainability-vol.pdf")
```


None of the scenarios with an initial proportion of commercial volume of 20% are sustainable after the first logging cycle (Figure \@ref(fig:tsust-vs-prodi); Table \@ref(tab:sust-production)). The present logging practices in the Brazilian Amazon with a proportion of commercial species not exceeding 20%, a mean logging intensity of 15-20 m$^3$.ha$^{-1}$ and a rotation cycle of 35 years are therefore not sustainable as there will be a significant decrease in timber production at the second cycle (< 20 m$^3$.ha$^{-1}$). 

Scenarios with a proportion of commercial volume of 50% show longer duration of constant production. The most sustainable one with a constant production of timber during `r dftsust[logIntensity=="10 m3/ha" & logCycle=="60 yr" & omega0=="50%", "Potential"]` allows a logging intensity of 10 m$^3$.ha$^{-1}$ every 60 years (Table \@ref(tab:sust-production); Figure \@ref(fig:tsust-vs-prodi)). Scenarios with a proportion of 90% of commercial species are among the most sustainable.

Only 4 out of all 27 scenarios have a median duration of constant production greater or equal to 500 years which can be considered as sustainable *sensu stricto*. Three of these scenarios have an initial proportion of commercial volume of 90%, and two correspond to low intensity logging (10 m$^3$.ha$^{-1}$) with a rotation cycle of 60 years (Table \@ref(tab:sust-production)). Logging regulations in the Brazilian Amazon include minimum logging cycles of 35 years. Under these regulations, the duration of constant production is `r dftsust[omega0=="50%"&logCycle=="35 yr"&logIntensity=="10 m3/ha"][, "Potential"]` if 50% of the volume is commercial, and `r dftsust[omega0=="90%"&logCycle=="35 yr"&logIntensity=="10 m3/ha"][, "Potential"]` if 90% of the volume is commercial (i.e. a true sustainable production). However these scenarios imply low logging intensities of 10 m$^3$.ha$^{-1}$ and a higher proportion of commercial timber species than in the present. According to our simulations and considering the 95% credibility interval, the most sustainable scenario is the one with an initial proportion of 90% of commercial species allowing a logging intensity of 10 m$^3$.ha$^{-1}$ every 60 years (Table \@ref(tab:sust-production)).


The current timber production in the Amazon is around 11 Mm$^3$ per year [@Vidal2020;@SFB2019]. Current concessions are not able to satisfy this demand alone for more than one cutting cycle (Figure  \@ref(fig:tsust-vs-prodi)). The maximum production under the current concession areas is `r signif(dfsave[logRules=="HS"&ppf=="Current concessions"&omega0=="omega0 = 90%", "med_prodi"], 3)/1e6` Mm$^3$ only during the first harvesting and a proportion of commercial species of 90%. Under the present practices (20 m$^3$.ha$^{-1}$) every 35 years and 20% of commercial species), the annual production during the first harvest reach `r signif(dfsave[logRules=="MM"&ppf=="Current concessions"&omega0== "omega0 = 20%","med_prodi"], 2)/1e6` Mm$^3$ only. Finally, under current concession area four scenarios can be considered as sustainable (duration of constant production > 500 years). The highest sustainable production is recorded under an extraction rate of 20 m$^3$.ha$^{-1}$ every 60 years with a 90% proportion of commercial species (Figure \@ref(fig:tsust-vs-prodi)).

When considering all potential concession areas (35 Mha), the annual production of 11 Mm$^3$.yr$^{-1}$ could be sustained at best during `r dftsust[omega0=="90%"& logCycle=="35 yr"&logIntensity=="20 m3/ha"][, "Potential"]` following a scenario with 90% of commercial species, a logging intensity of 20 m$^3$.ha$^{-1}$  and a rotation cycle of 35 years (Figure \@ref(fig:tsust-vs-prodi); Table 1). There are two others scenarios close to 11 Mm$^3$ with a sustained duration of 250 years (Figure \@ref(fig:tsust-vs-prodi)), with a logging intensity of 10 and 30 m$^3$.ha$^{-1}$ and a logging cycle length of 20 and 60 years, respectively. 


# Discussion

- methods: optimistic hypotheses -> to be considered when interpreting results -> Camille

- According to the results of our simulations, several challenges must be faced to reach sustained timber production within the concession systems in the Brazilian Amazon. First, the proportion of commercial species must increase considerably at least to 50% but ideally to 90%. Second, the logging intensity must be reduced to 10 m$^3$.ha$^{-1}$ and the rotation cycle to 60 years. Third, the area of current concessions is obviously insufficient to reach an annual production of 11 Mm$^3$ and it is urgent to increase drastically concession area to its maximum potential of 35 Mha. However, the potential concession area of 35 Mha  show strong limitation to ensure a annual sustainable production of 11 Mm$^3$ on a long term basis as our simulations suggest that this production can be sustained for at best 170 years with a 90% proportion of commercial species. 

- This study confirms that the present regulation fixing a rotation cycle of 35 years  cannot ensure a long term sustainable timber production. Although previous studies already demonstrated the need to change the present extraction rates (Sist and Ferreira 2007, Putz et al. 2008, other refs) our paper is the first one suggesting and assessing different options of logging practices. The main challenge to move towards more sustainable practices is undoubtedly to increase the proportion of commercial species from presently 20 % to 90%. This must involve drastic changes at all level but particularly in the industrial and market sectors. At present, the wood transformation industry is still technically very poorly developed and very rudimentary (refs?). Less than 30% of the logs entering the saw mill is transformed, the remaining 70% are wasted and burned. However, the lack of wood transformation technology is certainly linked to the lack of market demand on diversified wood products in the region. To be developed (Jack ? Edson ?)

- increasing the area of concessions: current trends, challenges and opportunities

- initial proportion of commercial species: what effect could harvesting 90% of all trees > 50 cm DBH could have on the profitability of selective logging?

- initiating a forest transition -> new ways of producing timber: plantations, active restoration, silviculture

- what effects could silviculture have on sustainability? -> needs to be investigated (introduce next paper)



# References
