---
title: "Sustainable logging - main document"
output: 
  html_document:
    theme: journal
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, 
                      root.dir = rprojroot::find_rstudio_root_file())
# load libraries
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
```

# Exploring timber volume trajectories

We will first test 5 scenarios and look at the predicted trajectories. The location stays the same (here we use Manaus coordinates in central Amazonia).

Four input variables differ among scenarios: 

- the logging intensity (in m$^3$ha$^{-1}$)

- the length of the logging cycle (in years)

- $\omega 0$: the pre-logging proportion of commercial timber 

- $dti$: the pre-logging disturbance factor (between 0 and 1): when $dti=0$, the site has suffered no human disturbance previous to logging; if $dti=1$, the site has been completely disturbed and clearcut. 

```{r}
timeLength = 1000
logIntensity = rep(10, 5); logIntensity[2] = 30
logCycle = rep(35, 5); logCycle[3] = 65
omega0 = rep(0.70, 5); omega0[4] = 0.90
longitude = rep(-60.6, 5) ## location: Manaus
latitude = rep(-3.3, 5)
dti = rep(0.10, 5); dti[5] = 0.40
tabScenario = data.table(scenario = c("default", "highIntensity", "longCycle", "divCommSpecies", "prevDisturb"), 
                         logIntensity, logCycle, omega0, dti)
```

## Without uncertainties

The following table summarizes the 5 scenarios and the values of each input variable. 

```{r}
tabScenario %>%
  kable() %>%
  kable_styling()
```

The following figure shows the predicted trajectory of timber volume stocks for each of the 5 scenarios. 

```{r illustr_traj, fig.height=4, fig.width=10}

source("functions/timbRecovery.R")

recov = timbRecovery(timeLength, logIntensity, logCycle, omega0, longitude, latitude, dti)

dfrecov = do.call(data.table, recov[[1]])
colnames(dfrecov) = tabScenario$scenario
dfrecov$t = 1:timeLength
dfrecov = melt(dfrecov, id.vars = "t", value.name = "volume", variable.name = "scenario")

ggplot(dfrecov, aes(x=t, y = volume, color = scenario)) + geom_line() + xlab("Time since first logging event (yr)") + ylab("Commercial timber volume (m3/ha)") 
```

The following figure shows the predicted trajectory of timber volume stocks for each of the 5 scenarios. 

```{r illustr_vextReal, fig.height=2, fig.width=10}
vextReal = data.table(t(recov[[2]]))
colnames(vextReal) = tabScenario$scenario
vextReal$ncycle = 1:nrow(vextReal)
vextReal = melt(vextReal, id.vars = "ncycle", value.name = "intensity", variable.name = "scenario")

ggplot(vextReal, aes(x= ncycle, y = intensity, colour = scenario)) + geom_line() + facet_wrap(~scenario, nrow = 1) + 
  xlab("Number of cutting cycles") + ylab("Volume extracted")
```

Only the "longCycle" scenario maintains high levels of timber volume and a constant production after 1000 years; it can be considered sustainable. 


## With uncertainties

The following figure shows the predicted trajectory of timber volume stocks for each of the 5 scenarios. 

```{r illustr_traj_uncert, fig.height=4, fig.width=10}

source("functions/timbRecoveryUncert.R")

ptm <- proc.time()
recov = timbRecoveryUncert(timeLength, logIntensity, logCycle, omega0, longitude, latitude, dti)
proc.time() - ptm

dfrecov = recov[[1]]
dfrecov$scenario = tabScenario$scenario[dfrecov$site]

ggplot(dfrecov, aes(x=t, y = med, color = scenario, ymin = inf, ymax = sup)) + 
  geom_ribbon(aes(fill = scenario), colour = NA, alpha = 0.1) + geom_line() + theme(legend.position = "none") +
  xlab("Time since first logging event (yr)") + ylab("Commercial timber volume (m3/ha)")  + facet_wrap(~scenario)

```

The following figure shows the predicted trajectory of timber volume stocks for each of the 5 scenarios. 

```{r illustr_vextReal_uncert, fig.height=2, fig.width=10}
vextReal = data.table(recov[[2]])
vextReal$scenario = tabScenario$scenario[vextReal$site]

ggplot(vextReal, aes(x = ncycle, y = med, color = scenario, ymin = inf, ymax = sup)) + 
  geom_ribbon(aes(fill = scenario), colour = NA, alpha = 0.1) + geom_line() + theme(legend.position = "none") +
  facet_wrap(~scenario, nrow = 1) + xlab("Number of cutting cycles") + ylab("Volume extracted")
```


For now we can say that logging is sustainable if the production stays constant during the first 30 cycles. 

