---
title: "Silvivulture"
author: "Camille Piponiot"
date: "2/12/2021"
output: word_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, results='hide', 
                      root.dir = rprojroot::find_rstudio_root_file())
# load libraries
packages_needed <- c("knitr", "data.table", "ggplot2", "parallel", "sp", "truncnorm")
packages_to_install <- packages_needed[!( packages_needed %in% rownames(installed.packages()))]
if (length(packages_to_install) > 0)
  install.packages(packages_to_install)
lapply(packages_needed, require, character.only = TRUE)

## source all functions
file.sources = list.files(path = "R/", pattern="*.R", full.names = TRUE)
sapply(file.sources, source, .GlobalEnv)

if (!dir.exists("cache"))
  dir.create("cache")
```

### Introducing a new module: silvicultural treatments

The VDDE model as presented in @Piponiot2019a represents the most widespread type of selective logging in the tropics, which consists of selectively harvesting some trees and leaving the forest to recover with no additional human intervention. Some treatments can however be applied to the stand before, during or after the harvest to increase post-logging timber recovery. These silvicultural treatments usually consist of killing competitor trees or removing overtopping lianas from future crop trees [REF]. Removing competitor trees can have several disadvantages, such as emitting large amounts of carbon and killing potential future commercial species. Liana removal has been shown to be a cost-effective silvicultural treatment with long-lasting effects: it increased by xx-xx% the growth of targeted trees, and its effect can last for xx years [REFS].

In this study we added a module to the VDDE model that simulates the effect of silvicultural treatments such as liana removal. To do so, we increase the growth of commercial volume by a coefficient $silv$, the effect of silviculture, defined as a decreasing function of recovery time $t_{rec}$ (i.e. time since the last logging event, when the treatment is applied). This function is an adaptation of the logistic function (see equation xx and Figure \@ref(fig:illustr-silviculture)). The codes are provided in the appendix xx.

$$silv(t_{rec}) = Smax\cdot\left(1 - \frac{1}{exp(\lambda \cdot (t_{80}-t_{rec})+log(4)) + 1} \right)$$

where $t_{80}$ is the time when the function reaches 80% of its initial value $Smax$ and $\lambda$ controls the steepness of the decrease. In this study we drew (positive) parameter values from the following lognormal distributions: $Smax \sim log\mathcal{N} (log(0.5), \frac{log(1)-log(0.5)}{2})$, $t_{80} \sim log\mathcal{N} (log(20), \frac{log(50)-log(20)}{2})$ and $\lambda \sim log\mathcal{N} (log(0.25), \frac{log(1)-log(0.25)}{2})$.
Under those distribution, the median of parameters $Smax$, $t_{80}$ and $\lambda$ is respectively 0.5, 20 years and 0.25 yr$^{-1}$, and their 97.5$^{th}$ percentile is respectively 1, 50 yr and 1 yr$^{-1}$. The values of the distributions were chosen based on the literature [REFS], and with large standard deviation to represent the uncertainty over these estimates.

> idea: lambda and t80 can also be randomly drawn

> data: @Roopsind2018: liana + girdling, +150% increase in volume recovery after 20 years

```{r illustr-silviculture, fig.width = 8, fig.height=4, fig.caption = "Silviculture effect as a function of recovery time."}
## parameters
tmax = 50
ag = 5
bg = 0.01
am = 4
bm = 0.005
theta = 0.002
intpR = 5
slopepR = -1.5
om0 = 0.8
lbd = 0.25 # steepness of the decrease

# test different silvilcultural effects
## increase in growth of commercial trees
silv_effect = list(1 + 0*(1 - 1 / (exp(lbd*(0-(1:tmax))+log(4)) + 1)), ## no effect
                   1 + 0.2*(1 - 1 / (exp(lbd*(10-(1:tmax))+log(4)) + 1)), ## 20% - 10 years
                   1 + 0.2*(1 - 1 / (exp(lbd*(20-(1:tmax))+log(4)) + 1)), ## 20% - 20 years
                   1 + 0.5*(1 - 1 / (exp(lbd*(10-(1:tmax))+log(4)) + 1)), ## 50% - 10 years
                   1 + 0.5*(1 - 1 / (exp(lbd*(20-(1:tmax))+log(4)) + 1))) ## 50% - 20 years

names(silv_effect) <- c("none", "20% - 10 years", "20% - 20 years", "50% - 10 years", "50% - 20 years")

full <- FALSE
if (full){
  simulations <- lapply(silv_effect, function(K) {
    # initialization
    V = c(100)
    om = c(0.2)
    vtest = volume(t = 1:500, ag = ag, am = am, bg = bg, bm = bm, th = theta)
    t0_test = which.min(abs(vtest-100))
    for (t in 1:(tmax-1)) {
      pR = 1 / (1 + exp(-(intpR + slopepR * log(V[t]))))
      eta = pR * (om0*K[t] + 1 - om0) + (1 - pR) * (om[t]*K[t] + 1 - om[t])
      g = (ag*(1-exp(-bg*(t0_test+t))) - theta*V[t])*eta
      m = am*(1-exp(-bm*(t0_test+t)))
      V[t+1] = V[t] + g - m
      om[t+1] = min((V[t]*om[t] + (ag*(1-exp(-bg*(t0_test+t))) - theta*V[t]) * K[t] * (om0*pR + om[t]*(1-pR)) - m*om[t] ) / V[t+1], 1)
    }
    return(data.frame(t = 1:tmax, silv = (K-1)*100, vol = V, omega = om*100))
  })

  simulations <- rbindlist(simulations, idcol = "treatment")
  simulations[, vcom := vol*omega/100]
  dfsim <- melt(simulations, measure.vars = c("silv", "vol", "omega", "vcom"))
  levels(dfsim$variable) <- c("Effect of silvicultural treatments\non commercial volume growth (%)",
                              "Total volume (m3/ha)", "Proportion of commercial volume (%)", "Commercial volume (m3/ha)")

  ## graph
  library(ggplot2)
  ggplot(dfsim, aes(x = t, y = value, color = treatment)) +
    geom_line() +
    facet_wrap(~ variable, scales = "free") +
    theme_bw() +
    expand_limits(x = 0, y = 0) +
    scale_x_continuous(expand = c(0, 0)) +
    labs(x = "Recovery time (yr)", y = "")
} else {
  df <- data.frame(treatment = rep(names(silv_effect), each = tmax),
                   trec = rep(1:tmax, length(silv_effect)),
                   silv = (unlist(silv_effect)-1)*100)
  # dflines <-
  ggplot(df, aes(x = trec, y = silv, color = treatment)) +
    geom_line()+
    theme_classic() +
    annotate(geom = "segment", lty = 2,
             x = c(0, 0, 10 , 20),
             xend = c(20 , 20, 10, 20),
             y = c(40, 16, 0, 0),
             yend = c(40, 16, 40, 40)) +
    scale_x_continuous(expand = c(0, 0)) +
    labs(x = "Recovery time (yr)", y = "Increase in commercial volume growth (%)")
}

```

